Tail associations in ecological variables and their impact on extinction risk

### 2020-01-21 MainText.Rmd
minor stuff during first half of a read through

<!--objectives, last para of intro-->
In this paper, we will explore the potential for asymmetric tail associations to 
influence extinction risk in populations and metapopulations. Specifically,
we will address the following sets of questions. Q1) Can the nature of tail associations
between vital rates in an age-structured population model influence
extinction risk for the model, and to what extent? Q2) Does the nature
of tail associations between measurements of an environmental variable in different
locations influence the extinction risk of a density-dependent metapopulation model
that depends on the environmental variable? How does the nature of the density
dependence mediate this influence? Will density dependence reverse or reproduce
the earlier result [@Ghosh_copula], using non-density-dependent models, that 
left-tail dependence in environmental variables accentuates and right-tail dependence 
mitigates extinction risks?
For Q1, we simulate a simple 2-stage model. We hypothesize that left-tail dependence
between juvenile survival and adult fecundity rates increases extinction risk
relative to right-tail dependence, even when the overall correlation is kept fixed. 
For Q2, we simulate spatial extensions of several simple and commonly
used population models. We intentionally use a simulation approach and simple models
for both Q1 and Q2 instead of attempting mathematical 
analysis and/or using models that are more realistic, 
because our intention here is to provide an initial exploration only of 
the potential mechanisms outlined above. 
By so doing we hope to open these potentially practically important questions to 
wider examination by other researchers, using a variety of models and perspectives. 
Though tail associations are used
to assess risks in fields such as finance and hydrology 
[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
knowledge asymmetric tail associations have rarely been assessed for their
importance for extinction risks of populations and species (but see @Ghosh_copula),
and our essential goal is to advance the idea that tail associations may
also be important in this context. 
Our questions are also an important step for understanding
the potential influence of ECEs on extinction risks if ECEs are related to
tail associations in environmental variables.

>>>>

<!--objectives, last para of intro-->
In this paper, we will explore the potential for asymmetric tail associations to 
influence extinction risk in populations and metapopulations. Specifically,
we will address the following sets of questions. Q1) Can the nature of tail associations
between vital rates in an age-structured population model influence
extinction risk for the model, and to what extent? Q2) Does the nature
of tail associations between measurements of an environmental variable made in different
locations influence the extinction risk of a density-dependent metapopulation model
that depends on the environmental variable? How does the nature of the density
dependence mediate this influence? Will density dependence reverse or reproduce
the earlier result [@Ghosh_copula], using non-density-dependent models, that 
left-tail dependence in environmental variables accentuates and right-tail dependence 
mitigates extinction risks?
For Q1, we simulate a simple 2-stage model. We hypothesize that left-tail dependence
between juvenile survival and adult fecundity rates increases extinction risk
relative to right-tail dependence, even when the overall correlation is kept fixed. 
For Q2, we simulate spatial extensions of several simple and commonly
used population models. We intentionally use a simulation approach and simple models
for both Q1 and Q2 instead of attempting mathematical 
analysis and/or using models that are more realistic and complex, 
because our intention here is to provide an initial exploration only of 
the potential mechanisms outlined above. 
By so doing we hope to open these potentially practically important questions to 
wider examination by other researchers, using a variety of models and perspectives. 
Though tail associations are used
to assess risks in fields such as finance and hydrology 
[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
knowledge asymmetric tail associations have rarely been assessed for their
importance for extinction risks of populations and species (but see @Ghosh_copula),
and our essential goal is to advance the idea that tail associations may
also be important in this context. 
Our questions are also an important step for understanding
the potential influence of ECEs on extinction risks if ECEs are related to
tail associations in environmental variables.
